---
title: "ProjectWork"
author: "Maame Efua Boham"
date: "October 27, 2017"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

**Is there a relationship between attitude of parents and use of birth control?**

**Is there a relationship between timing of first sexual intercourse and perceptions of parental liberality?**

## 1. Load data sets and libraries
```{r}
load("C:/Users/Boham_/Desktop/School/Statistics/Project/datasets/addhealth_W1.RData")
library(descr)
```

## 2. Create data subset
```{r}
vars = c("AID","BIO_SEX","age", "H1CO2Y", "H1CO3", "H1CO6", "H1CO13", "H1PA1", "H1PA3", "H1PA4", "H1PA6")

temp = AddHealth[vars]

myData = subset(AddHealth,age > 15 & H1CO1==1 )

```

## 3. Remove error codes and other data management
```{r}
myData$BIO_SEX[myData$BIO_SEX == 6 ] <- NA



myData$H1CO2Y[myData$H1CO2Y == 96 | myData$H1CO2Y == 98 | myData$H1COY == 99] <- NA

myData$H1CO2Y[myData$H1CO2Y == 97] <- NA



myData$H1CO3[myData$H1CO3 == 6 | myData$H1CO3 == 8] <- NA

myData$H1CO3[myData$H1CO3 == 7] <-  NA



myData$H1CO6[myData$H1CO6 == 6 | myData$H1CO6 == 8] <- NA

myData$H1CO6[myData$H1CO6 == 7] <-  NA



myData$H1CO13[myData$H1CO13 == 6 | myData$H1CO13 == 8] <- NA

myData$H1CO13[myData$H1CO13 == 7] <-  NA



myData$H1PA1[myData$H1PA1 == 6 | myData$H1PA1 == 8 | myData$H1PA1 == 9] <- NA

myData$H1PA1[myData$H1PA1 == 7] <-  NA



myData$H1PA3[myData$H1PA3 == 6 | myData$H1PA3 == 8 | myData$H1PA3 == 9] <- NA

myData$H1PA3[myData$H1PA3 == 7] <-  NA



myData$H1PA4[myData$H1PA4 == 6 | myData$H1PA4 == 8 | myData$H1PA4 == 9] <- NA

myData$H1PA4[myData$H1PA4 == 7] <-  NA



myData$H1PA6[myData$H1PA6 == 6 | myData$H1PA6 == 8 | myData$H1PA6 == 9] <- NA

myData$H1PA6[myData$H1PA6 == 7] <-  NA


myData$MAttBC = rep(NA,nrow(myData))
myData$MAttBC[myData$H1PA3 == 1 | myData$H1PA3 == 2] = 1
myData$MAttBC[myData$H1PA3 == 3] = 2
myData$MAttBC[myData$H1PA3 == 4 | myData$H1PA3 == 5] = 3


levels(myData$MAttBC) = c("Disapprove","Neutral","Approve")

myData$PAttBC = rep(NA,nrow(myData))
myData$PAttBC[myData$H1PA6 == 1 | myData$H1PA6 == 2] = 1
myData$PAttBC[myData$H1PA6 == 3] = 2
myData$PAttBC[myData$H1PA6 == 4 | myData$H1PA6 == 5] = 3


levels(myData$PAttBC) = c("Disapprove","Neutral","Approve")



```

## 4. Univariate graphing for quantitative variables
```{r}
hist(myData$age, main = "Wave 1 Age Distribution", xlab = "age")
summary(myData$age)

hist(myData$H1CO2Y, main = "Year of First Sex Distribution", xlab = "year")
summary(myData$H1CO2Y)


```

## 5. Univariate graphing for categorical  variables
```{r}

freq(droplevels(myData$BIO_SEX))


freq(droplevels(myData$BIO_SEX),y.axis="percent",main="Gender Distribution in AddHealth", names = c("male","female"),ylab="percent")


freq(droplevels(myData$H1CO3),y.axis="percent",main="Birth Control Use Distribution in AddHealth", names = c("no","yes"),ylab="percent")

freq(droplevels(myData$H1CO6),y.axis="percent",main="Distribution of Birth Control Use during Most Recent Sex", names = c("no","yes"),ylab="percent")

freq(droplevels(myData$H1CO13),y.axis="percent",main="Regular Birth Control Pill Use Distribution", names = c("no","yes"),ylab="percent")


freq(droplevels(myData$H1PA1),y.axis="percent",main="Parental Attitude Towards Sex(Mom) in AddHealth", names = c("str. disapprove","disapprove","neutral","approve","strongly approve"),ylab="percent")

freq(droplevels(myData$H1PA3),y.axis="percent",main="Parental Attitude Towards Birth Control(Mom)", names = c("str. disapprove","disapprove","neutral","approve","strongly approve"),ylab="percent")

freq(droplevels(myData$H1PA4),y.axis="percent",main="Parental Attitude Towards Sex(Dad) in AddHealth", names = c("str. disapprove","disapprove","neutral","approve","strongly approve"),ylab="percent")

freq(droplevels(myData$H1PA6),y.axis="percent",main="Parental Attitude Towards Birth Control(Dad)", names = c("str. disapprove","disapprove","neutral","approve","strongly approve"),ylab="percent")




```

## 6. Bivariate output and graphing
```{r}

t1<-table(myData$H1CO6,droplevels(myData$BIO_SEX));t1

t1_col_prop<-round(prop.table(t1, 2),3); t1_col_prop

barplot(t1_col_prop[2, ], main="Birth Control Used During Most Recent Sex BY Gender", ylab="Proportion", names=c("male","female"))

mom_table <- table (myData$H1CO6, myData$MAttBC);mom_table
mom_table_colProp <- prop.table(mom_table, 2);mom_table_colProp
barplot (mom_table_colProp [2, ], main="BIRTH CONTROL USE BY MOTHER'S ATTITUDE TOWARDS BIRTH CONTROL", ylab="Proportion of adolescents who used birth control during most recent sex", names=c("disapproved","neutral","approved"), col = "purple4")

dad_table <- table (myData$H1CO6, myData$PAttBC);dad_table
dad_table_colProp <- prop.table(dad_table, 2);dad_table_colProp
barplot (dad_table_colProp [2, ], main="'BIRTH CONTROL USE BY FATHER'S ATTITUDE TOWARDS BIRTH CONTROL", ylab="Proportion of adolescents who used birth control during most recent sex", names=c("disapproved","neutral","approved"),col = "purple4")



```
 **Is attitude of mother towards birth control( MAttBC - categorical explanatory variable) associated with use of birth control, and is the association different for male and female adolescents(BIO_SEX - moderating categorical variable)?**
 
 **Is attitude of father towards birth control( PAttBC - categorical explanatory variable) associated with use of birth control(H1CO3 - categorical response variable ), and is the association different for male and female adolescents(BIO_SEX - moderating categorical variable)?**
 
 
 
*age - change in year versus year of first sex
 





## 7. Bivariate analysis
```{r}

myData$MAttBC<-as.factor(myData$MAttBC)
myData$H1CO6<-as.factor(myData$H1CO6)

chisq_results <- chisq.test(myData$H1CO6, myData$MAttBC)
chisq_results 

chisq_results$observed
chisq_results$expected
chisq_results$residuals


myData$PAttBC<-as.factor(myData$PAttBC)
myData$H1CO6<-as.factor(myData$H1CO6)

chisq_results <- chisq.test(myData$H1CO6, myData$PAttBC)
chisq_results 

chisq_results$observed
chisq_results$expected
chisq_results$residuals


mySubset_1_2 <- subset(myData, MAttBC == 1 | MAttBC == 2)
chisq_1_2 <- chisq.test(mySubset_1_2$H1CO6, mySubset_1_2$H1CO6)
chisq_1_2

mySubset_1_3 <- subset(myData, MAttBC == 1 | MAttBC == 3)
chisq_1_3 <- chisq.test(mySubset_1_3$H1CO6, mySubset_1_3$H1CO6)
chisq_1_3

mySubset_2_3 <- subset(myData, MAttBC == 2 | MAttBC == 3)
chisq_2_3 <- chisq.test(mySubset_2_3$H1CO6, mySubset_2_3$H1CO6)
chisq_2_3


```

## 8. Moderation
```{r}

freq(myData$MAttBC)
myData$H1CO6<-as.numeric(myData$H1CO6)
tbl <- ftable(by(myData$H1CO6, list(droplevels(myData$BIO_SEX),myData$MAttBC),sum, na.rm = TRUE))
tbl

p_tbl <- prop.table(tbl, 2)          
p_tbl                                               
barplot(p_tbl, beside = TRUE, main ="BIRTH CONTROL USE AND MOTHER'S ATTITUDE TOWARDS BIRTH CONTROL BY GENDER", names.arg = c("Disapproved","Neutral","Approved"), ylab="Proportion of adolescents who used birth control during most recent sex", xlab="Attitude of Mother towards Birth Control",col=c("purple2","purple4"))
legend("topleft", legend=c("Male","Female"),fill=c("purple2","purple4"))

freq(myData$PAttBC)
myData$H1CO6<-as.numeric(myData$H1CO6)
tbl <- ftable(by(myData$H1CO6, list(droplevels(myData$BIO_SEX),myData$PAttBC),sum, na.rm = TRUE))
tbl

p_tbl <- prop.table(tbl, 2)          
p_tbl                                               
barplot(p_tbl, beside = TRUE, main ="BIRTH CONTROL USE AND FATHER'S ATTITUDE TOWARDS BIRTH CONTROL BY GENDER", names.arg = c("Disapproved","Neutral","Approved"), ylab="Proportion of adolescents who used birth control during most recent sex", xlab="Attitude of Father towards Birth Control",col=c("purple2","purple4"))
legend("topleft", legend=c("Male","Female"),fill=c("purple2","purple4"))


by(myData, myData$BIO_SEX, function(x) 
             list( chisq.test(x$H1CO6, x$PAttBC), 
             chisq.test(x$H1CO6, x$PAttBC)$observed, 
             prop.table(chisq.test(x$H1CO6, x$PAttBC)$observed, 2)))

by(myData, myData$BIO_SEX, function(x) 
             list( chisq.test(x$H1CO6, x$MAttBC), 
             chisq.test(x$H1CO6, x$MAttBC)$observed, 
             prop.table(chisq.test(x$H1CO6, x$MAttBC)$observed, 2)))


```

## 9. Save and quit
```{r}

```


 